---
title: "simpleRegress"
author: "Edward Gomez"
date: "Fall 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

This is weight predicted by age. The adjusted r-squared value is 0.5526. The residual standard error is 81.47.

```{r}
simp <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

This is weight predicted by head length. The adjusted r-squared value is 0.69. The residual standard error is 67.81

```{r}
simp <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

This is weight predicted by head width. The adjusted r-squared value is 0.6064. The residual standard error is 76.41.

```{r}
simp <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp)
summary.lm(simp)
```

This is weight predicted by length. The adjusted r-squared value is 0.7423. The residual standard error is 61.83.

```{r}
simp <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

This is weight predicted by chest. The adjusted r-squared value is 0.9262. The residual standard error is 33.08.

DISCUSSION AND CONCLUSION

Ranking for weight prediction from worst to best
1.Age ARS:0.5526 RSE:81.47
2.Head width ARS:0.6064 RSE:76.41
3.Head length ARS:0.69 RSE:67.81
4.Length ARS:0.7423 RSE:61.83
5.Neck ARS:0.87 RSE:43.91
6.Chest ARS:0.9262 RSE:33.08

This ranking shows that chest is the best metric for predicting weight.